Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.

### 2022-06-06 manuscript.Rmd
clarify terminology

First, the theory itself limits theoretical risk:
Some theories simply do not make risky predictions, and preregistration will not change that.
Consider the case of a researcher contemplating the relation between two sets of variables.
Suppose each set by itself is well studied, and strong theories tell the researcher how the variables within the set relate.
However, our imaginary researcher considers the relation between these two sets.
For lack of a better theory, they assume that some relation between any variables of the two sets exists.
This is not a risky prediction to make in psychology, even without statistical issues like alpha inflation [@orbenCrudReDefined2020].
However, we would consider it a success if the researcher would use the evidence to develop a more precise (and therefore risky) theory, e.g. by specifying which variables from one set relate to which variables from the other set, to what extent, in which direction, with which functional shape etc.
We will later show that preregistration increases the expected epistemic value of the further specified theory, though it remains low till substantiated by testing it again.
The point, however, is that we want to show that preregistration increases the expected epistemic value without regard to the theory being tested.

@@

Preregistration is often discussed in relation to statistical methods, so let us note some relations between $P(E|¬H)$ and statistical methods.
If you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, then the nill-hypothesis is $¬H$, and therefore $P(E|¬H)$ is equivalent to the type I error rate.
Researchers who choose a smaller type I error rate can hence be more sure of their results, if significant.
However, the research hypothesis is seldomly equal to the statistical null hypothesis.
We argue that $P(E|¬H)$ also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.

@@

The point is that the contribution of statistical methods to $P(E|¬H)$ can be formally assessed.
For many statistical models, their $P(E|¬H)$ can be analytically computed under some assumptions.
For those models or assumptions where this is impossible, one can employ Monte Carlo simulation to estimate the contribution to $P(E|¬H)$.
The precision with which researchers can discuss statistical contributions to $P(E|¬H)$ has lured the community concerned with research methods into ignoring other factors that are much more uncertain.
We can not hope to resolve this uncertainty; but we have to be aware of its implications.
These are relayed in the following.

>>>>

First, the theory itself limits theoretical risk:
Some theories simply do not make risky predictions, and preregistration will not change that.
Consider the case of a researcher contemplating the relation between two sets of variables.
Suppose each set by itself is well studied, and strong theories tell the researcher how the variables within the set relate.
However, our imaginary researcher considers the relation between these two sets.
For lack of a better theory, they assume that some relation between any variables of the two sets exists.
This is not a risky prediction to make in psychology, even without statistical issues like alpha inflation [@orbenCrudReDefined2020].
However, we would consider it a success if the researcher would use the evidence to develop a more precise (and therefore risky) theory, e.g. by specifying which variables from one set relate to which variables from the other set, to what extent, in which direction, with which functional shape etc.
We will later show that preregistration increases the belief one can stake in the further specified theory, though it remains low till substantiated by testing it again.
The point, however, is that we want to show that preregistration increases the expected epistemic value without regard to the theory being tested.

@@

Preregistration is often discussed in relation to statistical methods, so let us note some relations between $P(E|¬H)$ and statistical methods.
If you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, then the nill-hypothesis is $¬H$, and therefore $P(E|¬H)$ is equivalent to the type I error rate.
Researchers who choose a smaller type I error rate can hence be more sure of their results, if significant.
However, the research hypothesis is seldomly equal to the statistical null hypothesis.
We argue that $P(E|¬H)$ (and hence its inverse, theorethical risk) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.

@@

The point is that the contribution of statistical methods to $P(E|¬H)$ can be formally assessed.
For many statistical models it can be analytically computed under some assumptions.
For those models or assumptions where this is impossible, one can employ Monte Carlo simulation to estimate the contribution to $P(E|¬H)$.
The precision with which researchers can discuss statistical contributions to $P(E|¬H)$ has lured the community concerned with research methods into ignoring other factors that are much more uncertain.
We can not hope to resolve this uncertainty; but we have to be aware of its implications.
These are relayed in the following.
